Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2020-12-15 manuscript.tex
Add paragraph about sampling fraction [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/10491ddaa088c71ad76a54ca7570ec8360f4c7ae.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/10491ddaa088c71ad76a54ca7570ec8360f4c7ae/checks
https://github.com/habi/acinar-analysis-manuscript/runs/423768179

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681} and is under revision in the \href{https://journals.physiology.org/journal/ajplung}{\emph{American Journal of Physiology - Lung Cellular and Molecular Physiology}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/d6fc17bd66bdca5cbff30d53e1dbc02d5351609b...eff43d9d3566f72329894e11e468eccd5bac3323}{has been updated} since the first round of review.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/eff43d9d3566f72329894e11e468eccd5bac3323/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/eff43d9d3566f72329894e11e468eccd5bac3323}{habi/acinar-analysis-manuscript@eff43d9}
on December 15, 2020.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(eff43d9)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(eff43d9)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Switzerland
\end{itemize}

@@

On single two-dimensional sections---either histological or virtual from tomographic imaging---it is not possible to unambiguously detect which parts of the airways and alveoli are connected three-dimensionally.
To detect and delineate individual acini the full three-dimensional information is needed.
Historically, single or very few acini were delineated and extracted from manually traced serial sections (\protect\hyperlink{ref-c3aIB7f9}{28}, \protect\hyperlink{ref-CPhMy3Ie}{50}) or sectioned silicone casts (\protect\hyperlink{ref-1GJECnKBu}{30}) with considerable manual work.
Non-destructive three-dimensional imaging is best suited to acquire data sets that can be used to easily detect, delineate, and study large amounts of single acini.
This has been shown by Vasilescu et al.~(\protect\hyperlink{ref-C6OaY84D}{48}).
In their study, they reconstructed 22 mouse acini from tomographic data sets of four 12 weeks old mice (at least postnatal day 84, fully developed lungs) and compared those with scanning electron microscope images of acinar silicone rubber casts.
Similarly, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) analyzed the volume of 32 mouse acini from six animals, as well as the branching pattern of their internal airways.
All prior work known to us has analyzed considerably fewer acini than the 268 acini presented here and focuses only on a time point, at which the lungs of the animals are already fully developed while our data spans the postnatal days 4 to 60.

@@

In the present study, a superset of the animals described by Haberthür et al.~(\protect\hyperlink{ref-7YLeeyu}{11}) were used.
Tschanz et al.~(\protect\hyperlink{ref-wnl86DEM}{46}) described a biphasic formation of new alveoli on the same set of animals.
The stereological analysis presented here represents a part of the 3R-initiative (replacement, reduction, and refinement) (\protect\hyperlink{ref-199ALtdJt}{53}) for the ethical use and the reduction of the number of animals sacrificed for science.

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung (\protect\hyperlink{ref-14OP85b2F}{2}) and that information drawn from one lobe is representative of the whole lung (\protect\hyperlink{ref-I9TmP6IU}{52}).
Furthermore, based on stereological data, Zeltner et al.~(\protect\hyperlink{ref-I9TmP6IU}{52}) were able to show that the lung parenchyma is quite homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, e.g.~between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy (\protect\hyperlink{ref-Zv1XcgAO}{4}).
A first analysis of regional difference shows similar results as mentioned above.

@@

As shown in a study by Osmanagic et al.~(\protect\hyperlink{ref-LzalnLBH}{29}) the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned my Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.

@@

It has been predicted by computational fluid dynamics simulations and by to-scale experiments that the amount and location of massless particle deposition, as well as ventilation, are depending on the size of the acini (\protect\hyperlink{ref-CVl41LwO}{8}, \protect\hyperlink{ref-1DP2FRUSZ}{16}, \protect\hyperlink{ref-RJ9GQcwl}{40}).
The predictions could be confirmed at least in one case, where mice received fluorescent particles during mechanical ventilation.
In 3D-visualizations the particles were predominately detected in the proximal half of the acini (\protect\hyperlink{ref-19jBv0ima}{51}).

@@

To our best knowledge, no in-depth characterization of the acinar properties as presented here---i.e.~spanning the lung development period---was available up to now.
Even though our results are not precisely accordant with data calculated from previous studies, they match well, found differences are discussed extensively.

@@

\leavevmode\hypertarget{ref-CPhMy3Ie}{}%
50. \textbf{Woodward JD}, \textbf{Maina JN}. A 3D digital reconstruction of the components of the gas exchange tissue of the lung of the muscovy duck, Cairina moschata. \emph{Journal of Anatomy} 206: 477--492, 2005. doi: \href{https://doi.org/10.1111/j.1469-7580.2005.00413.x}{10.1111/j.1469-7580.2005.00413.x}.

\leavevmode\hypertarget{ref-19jBv0ima}{}%
51. \textbf{Yang L}, \textbf{Feuchtinger A}, \textbf{Möller W}, \textbf{Ding Y}, \textbf{Kutschke D}, \textbf{Möller G}, \textbf{Schittny JC}, \textbf{Burgstaller G}, \textbf{Hofmann W}, \textbf{Stoeger T}, \textbf{Daniel Razansky}, \textbf{Walch A}, \textbf{Schmid O}. Three-Dimensional Quantitative Co-Mapping of Pulmonary Morphology and Nanoparticle Distribution with Cellular Resolution in Nondissected Murine Lungs..

\leavevmode\hypertarget{ref-I9TmP6IU}{}%
52. \textbf{Zeltner TB}, \textbf{Bertacchini M}, \textbf{Messerli A}, \textbf{Burri PH}. Morphometric Estimation of Regional Differences in the Rat Lung. \emph{Experimental Lung Research} 16: 145--158, 2009. doi: \href{https://doi.org/10.3109/01902149009087879}{10.3109/01902149009087879}.

\leavevmode\hypertarget{ref-199ALtdJt}{}%
53. Replacement, Reduction, Refinement {[}Online{]}. \emph{ALTEX} 19: 73--78, 2002. \url{https://www.altex.org/index.php/altex/article/view/1106} {[}15 Dec. 2020{]}.
\end{cslreferences}

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681} and is under revision in the \href{https://journals.physiology.org/journal/ajplung}{\emph{American Journal of Physiology - Lung Cellular and Molecular Physiology}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/d6fc17bd66bdca5cbff30d53e1dbc02d5351609b...10491ddaa088c71ad76a54ca7570ec8360f4c7ae}{has been updated} since the first round of review.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/10491ddaa088c71ad76a54ca7570ec8360f4c7ae/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/10491ddaa088c71ad76a54ca7570ec8360f4c7ae}{habi/acinar-analysis-manuscript@10491dd}
on December 15, 2020.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(10491dd)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(10491dd)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Switzerland
\end{itemize}

@@

On single two-dimensional sections---either histological or virtual from tomographic imaging---it is not possible to unambiguously detect which parts of the airways and alveoli are connected three-dimensionally.
To detect and delineate individual acini the full three-dimensional information is needed.
Historically, single or very few acini were delineated and extracted from manually traced serial sections (\protect\hyperlink{ref-c3aIB7f9}{28}, \protect\hyperlink{ref-CPhMy3Ie}{51}) or sectioned silicone casts (\protect\hyperlink{ref-1GJECnKBu}{30}) with considerable manual work.
Non-destructive three-dimensional imaging is best suited to acquire data sets that can be used to easily detect, delineate, and study large amounts of single acini.
This has been shown by Vasilescu et al.~(\protect\hyperlink{ref-C6OaY84D}{48}).
In their study, they reconstructed 22 mouse acini from tomographic data sets of four 12 weeks old mice (at least postnatal day 84, fully developed lungs) and compared those with scanning electron microscope images of acinar silicone rubber casts.
Similarly, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) analyzed the volume of 32 mouse acini from six animals, as well as the branching pattern of their internal airways.
All prior work known to us has analyzed considerably fewer acini than the 268 acini presented here and focuses only on a time point, at which the lungs of the animals are already fully developed while our data spans the postnatal days 4 to 60.

@@

In the present study, a superset of the animals described by Haberthür et al.~(\protect\hyperlink{ref-7YLeeyu}{11}) were used.
Tschanz et al.~(\protect\hyperlink{ref-wnl86DEM}{46}) described a biphasic formation of new alveoli on the same set of animals.
The stereological analysis presented here represents a part of the 3R-initiative (replacement, reduction, and refinement) (\protect\hyperlink{ref-199ALtdJt}{54}) for the ethical use and the reduction of the number of animals sacrificed for science.

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung (\protect\hyperlink{ref-14OP85b2F}{2}) and that information drawn from one lobe is representative of the whole lung (\protect\hyperlink{ref-I9TmP6IU}{53}).
Furthermore, based on stereological data, Zeltner et al.~(\protect\hyperlink{ref-I9TmP6IU}{53}) were able to show that the lung parenchyma is quite homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, e.g.~between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy (\protect\hyperlink{ref-Zv1XcgAO}{4}).
A first analysis of regional difference shows similar results as mentioned above.

@@

As shown in a study by Osmanagic et al.~(\protect\hyperlink{ref-LzalnLBH}{29}) the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned my Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.
The seemingly minuscule volume fraction of lung studied (around 1 \% or less) in this manuscript is---according to the stereological guidelines (\protect\hyperlink{ref-dNc8FfNn}{17})---ample volume to gain mathematically correct results if the stereology is carried out correctly as we have done to the best of our knowledge.
For human lungs for example a so-called sampling cascade is used to study details down to the thickness of the air-blood barrier (0.5 µm, see Figure 1 of (\protect\hyperlink{ref-1CLJ32t5m}{50}) for a visualization of this cascade).
The studied volume in this case is several decimal powers less than what is studied here and still mathematically sound and accepted.

@@

It has been predicted by computational fluid dynamics simulations and by to-scale experiments that the amount and location of massless particle deposition, as well as ventilation, are depending on the size of the acini (\protect\hyperlink{ref-CVl41LwO}{8}, \protect\hyperlink{ref-1DP2FRUSZ}{16}, \protect\hyperlink{ref-RJ9GQcwl}{40}).
The predictions could be confirmed at least in one case, where mice received fluorescent particles during mechanical ventilation.
In 3D-visualizations the particles were predominately detected in the proximal half of the acini (\protect\hyperlink{ref-19jBv0ima}{52}).

@@

To our best knowledge, no in-depth characterization of the acinar properties as presented here---i.e.~spanning the lung development period---was available up to now.
Even though our results are not in precise accordance with data calculated from previous studies, they match well relatively.

@@

\leavevmode\hypertarget{ref-1CLJ32t5m}{}%
50. \textbf{Weibel ER}, \textbf{Hsia CCW}, \textbf{Ochs M}. How much is there really? Why stereology is essential in lung morphometry. \emph{Journal of Applied Physiology} 102: 459--467, 2007. doi: \href{https://doi.org/10.1152/japplphysiol.00808.2006}{10.1152/japplphysiol.00808.2006}.

\leavevmode\hypertarget{ref-CPhMy3Ie}{}%
51. \textbf{Woodward JD}, \textbf{Maina JN}. A 3D digital reconstruction of the components of the gas exchange tissue of the lung of the muscovy duck, Cairina moschata. \emph{Journal of Anatomy} 206: 477--492, 2005. doi: \href{https://doi.org/10.1111/j.1469-7580.2005.00413.x}{10.1111/j.1469-7580.2005.00413.x}.

\leavevmode\hypertarget{ref-19jBv0ima}{}%
52. \textbf{Yang L}, \textbf{Feuchtinger A}, \textbf{Möller W}, \textbf{Ding Y}, \textbf{Kutschke D}, \textbf{Möller G}, \textbf{Schittny JC}, \textbf{Burgstaller G}, \textbf{Hofmann W}, \textbf{Stoeger T}, \textbf{Daniel Razansky}, \textbf{Walch A}, \textbf{Schmid O}. Three-Dimensional Quantitative Co-Mapping of Pulmonary Morphology and Nanoparticle Distribution with Cellular Resolution in Nondissected Murine Lungs..

\leavevmode\hypertarget{ref-I9TmP6IU}{}%
53. \textbf{Zeltner TB}, \textbf{Bertacchini M}, \textbf{Messerli A}, \textbf{Burri PH}. Morphometric Estimation of Regional Differences in the Rat Lung. \emph{Experimental Lung Research} 16: 145--158, 2009. doi: \href{https://doi.org/10.3109/01902149009087879}{10.3109/01902149009087879}.

\leavevmode\hypertarget{ref-199ALtdJt}{}%
54. Replacement, Reduction, Refinement {[}Online{]}. \emph{ALTEX} 19: 73--78, 2002. \url{https://www.altex.org/index.php/altex/article/view/1106} {[}15 Dec. 2020{]}.
\end{cslreferences}
